The Etiology of Cybersecurity

### 2021-03-01 reports/paper_0/main.tex
fix all oliviero's fixes

\fixnote{Oliviero}{Two general remarks. First, I'd maybe specify more what concretely A, B and F "look like", if that makes sense. How do we concretely specify a fact in the ABF framework? If I am not mistaken, RCC defines axiomatically how in general a mereotopology should be, and a *concrete* mereotopology corresponds to a particular system specification. Is this correct? If so, make explicit how one would encode this, if this make sense. At the very end you talk about a particular configuration being SAT w.r.t. the algebraic strucutre; is this algebraic structure, indeed, the system specification? You also sometimes speak about logic formulas; what is the connection between these logical formulas and regions? Is a region a "set" of logical formulas, or is it that a region is logically defined?

Secondly, and this is probably due to my inexperience with systems design (and ABF), it is not totally clear to me exactly what a belief is; more concretely, what is the difference between assertions and beliefs. Beliefs define the inputs and the outputs; assertions are what is exchanged between agents. Wouldn't the assertions already define the inputs and outputs? or are the beliefs the "expected" I/Os, while assertions are the "actual" I/Os? Indeed, a functional block is defined by its inputs/outputs. Then, if EQ(F,A) hold, aren't we already happy? Because, if all the inputs and outputs are compliant with your requirements, your system is functionally "correct", and behaves as you want. Am I mistaken? Perhaps there are some other aspects that I am not considering.}

\fixnote{Oliviero}{Question: the region F encodes the requirements. Then, what if we encode a requirement wrongly? or, what if we do not realise that some implict requirements are missing? Indeed, one might not see why integrity is a requirement until a breach happens. Is this within the scope of the theory? Finally; are we sure that all reasonable requirements can be encoded in the region F?}

@@

		belief (as epistemological

@@

beliefs, and the information or assertions it provides.

@@

As in \autocite{Steup2020epistemology}, the only objective of information is to
exchange beliefs between agents and, as defined in \autocite{Hintikka1993Information},

@@

beliefs to define the actual behavior of an agent or a system; i.e. we consider
behaviors as belief revision systems.

@@

\fixnote{Oliviero}{Is that actually what is happening? During this section III, we refer multiple times to engineering concepts such as system, functional specification and what not. We sometimes refer to the relationship between functional architecture and beliefs but that is not made precise. Maybe we should use only abstract terms, until the connection with those engineering aspects are introduced thoroughly? {\bf mr}: what I wanted to say here (and I probably failed) is that we define A, B, and F informally (and I changed abstractly with informally in the text) but we'll define them more precisely (and as engineering concept) later on. I modified the sentence to make this clearer but it seems to me to me that you're saying that the text is unnecessarily complicated by the mixed used of generic terms as beliefs and functional blocks/architecture. If that is the case we should definitely make the section clearer.}

@@

	\item We consider a set of axiomatic \emph{facts} 
		(definitory
		rules\footnote{Facts are definitory rule as they don't
		define how a real system is finally implemented but
		how it should be, trough a series of requirements. Those requirements 
		may not hold, through insecurity, in the implemented system.}\autocite{Hintikka1993Information})
		instead of considering the more general epistemic definition of
		knowledge. Specifically, facts describe: the functional
		architecture of each agent, and the physical/structural (HW/SW)
		architecture of each subsystem of agents and agent within a
		subsystem.

@@

\fixnote{Oliviero}{There are some parts I am perplexed about about this section from "There is no agreement..." up to here). It is unclear to me why we present and define the general concept of knowledge rather then the concept of fact, since we use mostly "facts" and as defined in [41] an agent is cmposed by assertions, beliefs and facts (and not knowledge). Indeed, it seems from your last bullet point that a fact is a particular case of knowledge (as you say, "more general"), but the exact definition of fact is not really clear from here, at least to me. Further; why are facts requirements? Intuitively, I would define a fact to be something that is true; requirements are something that we *want* to be true. I understood why we do so later on in the paper; we are later on interested, by RCC, in whether the beliefs of our agents are consistent (EQ) with the facts, that is, we want our system to behave consistently with the requirements. I would maybe make this explicit: we interpret requirements as facts, since we consider them to be the truth, towards which our agents must be compliant with (rather than: facts are requirements and not the actual truth). Another thing I found to be unclear is your definition of belief. You start by saying that we will define these things abstractly but you refer here to a very system-oriented definition of belief: inputs and outputs. Indeed, you say that beliefs define the behaviour, but the exact definition of belief was not super clear to me. {\bf mr}: I think you're right and the key points are (i) what's a fact? (ii) why stressing knowledge if abf is based on facts? (iii) what's the relation between facts and requirements? A fact is a statement which is taken as true. it's not necessarily true in general but the discussion is much complex (unfortunately). I wanted to add it to the paper but someone said that it may be tricky since those concepts have been used for 2k years at least. But, as an example, ``Socrates is a man'' is a fact (i.e. a true statement or, better, a statement taken as true), right? do we \emph{know} it? how? did we induce it from empirical evidences by testing every single man ever existed? I guess you see what I mean. So, in my limited view, facts are statement taken as true, knowledge can be expressed by true statements, and requirements are statement taken as true (by developers and everyone else). The link between those concepts seems rather strong but I think you are suggesting to simplify the text changing the focus on facts or detailing the relation between knowledge and facts. I chose the former and changed the text accordingly.}

@@

\fixnote{Oliviero}{Further: what is a region? Is it a "prime" concept, that is, a primitive concept? Or is it a set? if it is a primitive concept, then defining connectedness in terms of set operations such as $\subseteq$ is a little counterintuitive; if a region is a set (or some other well-known mathematical concept), then state so explicitly somewhere, imho (as I said at the beginning, explicitly describing how these ABF regions look like would make this much clearer, I think). {\bf mr} you're right. it's a primitive concept, I changed it in the tab and in the text.}

@@

Similarly to \autocite{Santaca2016abf}, we define agents as a meronomy (an
hierarchy of Part-Whole relations) over the constituents that we previously defined
(assertions, beliefs, and facts), based on a standard definition of mereology,
i.e. based on the definition of parthood relation between \emph{parts}.  Due to
the necessity of considering different relations between parts (as we'll show afterwards)
we extend the mereology to a
mereo-topology \autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
considering the relations in Table~\ref{tab:rcc358}.  For the sake of
readability, we use the term \emph{region} both to refer to a mereological part
and to a topological region.  Our aim is to create a meronomy (hierarchy of part-whole
relations)
instead of the taxonomies (categorization based on discrete sets) 
such as the one provided
in \autocite{NIST2020NVD,MITRE2020CVE} 
so that we don't need to rely on a scoring system (such as the 
CVSS) to assign a quantitative evaluation 
of the cybersecurity of each entry. Instead, we want a precise calculation
of the number of insecure configurations
\fixnote{Oliviero}{is this the same as a state? If so, maybe mention this in brackets in the definition. {\bf mr}: not necessarily but it will be in the next prototype and paper. For now, I'd keep it easy and general (plus, we'd need to introduce the concept of state oth.)} of a system to emerge from the mathematical
formulation of our cybersecurity hypothesis.

A mereotopology, as defined e.g. in \autocite{Rachavelpula2017mereotopology},
is a mathematical structure where the basic relation between regions
is the reflexive and symmetric relation \emph{Connects With} $\subseteq$,
that we use to order a universe of agents
$\agentuniverse$ (see in Table~\ref{tab:rcc358}).  We use the Region Connection
Calculus (RCC), as defined in \autocite{bennettLogics,improvingRCC}, to provide
an axiomatization of the mereo-topological concepts. In its broader definition,
the RCC theory is composed by eight axioms, and is known as RCC8. In the text,
for brevity, we will often focus only on RCC5 without loss of generality. Using
RCC5 instead of RCC8 prevents us from considering tangential connections
between spatial regions. However, tangential connections in RCC8 can be
considered as special cases of the more general spatial relations considered in
RCC5.

@@

\paragraph{Number of different configurations of a system}
The general formula to calculate the number of different types of agents is
$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$\fixnote{Oliviero}{What if our calculus uses relations of different arities? I guess we sum this up for all arities? {\bf mr}: I think so, and I see your point (we could generalize the formula) maybe it's not that important? not sure},
between $n$ different regions. Therefore, $r^{\binom{n}{k}}$
expresses the number of permutation of $r$
relations over ${\binom{n}{k}}$ elements with repetitions, 
with ${\binom{n}{k}}$ being the number of
$k$-ary combinations of $n$ regions.
In our case, $\binom{n}{k}=3$ since we consider $3$ regions 
($\assertionRegion,\beliefRegion,\factRegion$), and all the relations
considered in the RCC are binary.  Hence, using RCC5 (with five different
spatial relations) over three regions, we can theoretically define up to 125
different type of agents. However, only 54 of the 125 (as showed in
\autocite{improvingRCC}) combinations are topologically correct with respect to
the definition of the relations of RCC5. Generalizing to all the RCCs:

@@

\paragraph{\Rcc{$\assertionRegion$}{$\beliefRegion$} -- Collaboration} In order
to reason on the relation between assertions and beliefs we first need to
consider that, by definition, assertions are defined as transfer of information
between two agents, e.g., between two agents.  Therefore, as depicted in
Figure~\ref{fig:system-agent}, an agent has two main categories of assertions,
input and output assertions.  Given an agent $a$ and a collection of asserted
predicates $\Phi$
\fixnote{Oliviero}{If I am not mistaken, this is the first time we refer to assertions as concrete formulas/predicates. Perhaps we should make explicit before the "form" of such assertions? {\bf mr}: I'm not sure I got this point. Do you mean we should refer to a logic (e.g. FOL)? If so, we could but it has a cost (in terms of simplicity) and I'm not sure how to reuse that in the paper to justify that ``cost''},
the input assertions are those received by $a$ from an agent
$s$ acting as a sender, $\rassert{s}{a}{\Phi}$; similarly, output assertions
are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. With a slight abuse
of notation, in the text we simply write $\assertionRegion_{s\rightarrow a}$ when we abstract
away from the content of the assertion (i.e. $\Phi$).
We shall consider
two pairs of regions:%\footnote{``I am not asserting, as Lotze did, that a
%relation between $X$ and $Y$ consists of a quality in $X$ and a quality in $Y$
%-- a view which I regard as quite indefensible. -- I assert that a relation $Z$
%between $X$ and $Y$ involves the existence in $X$ of the quality ``having the
%relation $Z$ to $Y$'' so that a difference of relations always involves a
%difference in quality, and a change of relations always involves a change of
%quality.'' --  Ellis J. McTaggart, The Unreality of
%Time \autocite{Mctaggart1908unreality}}
\begin{itemize}
	\item \Rcc{$\assertionRegion_{s\rightarrow a}$}{$\beliefRegion$},
		where the relation between input-assertions and beliefs
		describes the soundness of the execution of the functional
		architecture w.r.t. input elicitation. With more details, the
		ideal specification of the functional architecture, along with
		the expected inputs, defines the functional behavior of an
		ideal system.  If all the inputs (assertions) are correctly
		handled in the functional specification (beliefs) the
		specification is complete.

@@

We are now in the position to define what a secure system is (with respect to
the $\abftheory$-framework) and, based on that definition, what the cybersecurity risk is and
how to quantify it in a risk matrix. The following definition holds for 
abstract systems defined in the $\abftheory$-framework but will be refined
for CPS afterwards in the paper.
\begin{definition}{\bf Cybersecurity of a System or an Agent --}\label{def:security}
	A secure system is a system where CSE-$1$, CSE-$2$, and CSE-$3$ holds for
	each agent composing the system.
\end{definition}
The ISO 31000 consider risk as the ``effect of uncertainty on objectives'' and
refers both to positive and negative consequences of uncertainty.
Accordingly, we consider risk as follows.
\begin{definition}{\bf Risk --}
	\fix{mr}{A risk is determined by the uncertainty related to the whole space of potential designs of a specification with respect to the
	$\abftheory$-framework.}
\end{definition}\fixnote{Oliviero}{iirc, this is the first time we speak about uncertainity. What do you mean exactly? Do you mean that we are uncertain about which state our CPS is in? Is it not knowing that we are in the secure state?}
The definition of Risk leads to the risk matrix in
Figure~\ref{fig:quantitative}, defined as follows.
\begin{definition}{\bf Risk Matrix --}
	The risk matrix, as summarized in Table~\ref{tab:5com}, is a function of the three relations
	$s=\langle\rcc(\factRegion,\beliefRegion),\rcc(\factRegion,\assertionRegion),\rcc(\beliefRegion,\assertionRegion)\rangle$,
	where the maximum risk is defined by the $DR$ relation between the
	three groups of regions, and the minimum risk by the $EQ$ relation over
	the same regions. In between the two extremes, the granularity of
	possible intermediate configuration is defined by the calculus used
	(RCC5 in our case).
\end{definition}\fixnote{Oliviero}{So, the granularity is just uniformly distributed between all possible configurations, right? Or is it the actual numeric value unimportant, as long as it "ordered" in the right way? {\bf mr}: right, the former}
While a risk matrix is often defined as a function of the likelihood and impact
of attacks (based on quantitative ad-hoc estimation of how likely it is that an
attacker will exploit one or more vulnerabilities and what is the magnitude of
the incidents produced by this exploitation), we suggest that cybersecurity
weaknesses are all equally likely to be exploited if there's a connection
between the weakness and the asset that an attacker wants to impact. When the
whole system is considered an asset, all weaknesses are equally likely to be
exploited.  Therefore, a risk matrix should capture the number of insecure
configurations of a system, rather than predicating over likelihoods of
weaknesses exploitation.

@@

So far, we have reasoned on systems in the abstract, considering
generic Multi-Agent System (MAS); mapping Epistemological concepts
to MAS. We now map MAS to CPS and show how this mapping allows us
to predict cybersecurity weaknesses in a system.\fixnote{Oliviero}{again, to me we actually referred quite a lot to systems and architectures.}

@@

As depicted in Figure~\ref{fig:mashyp}, we relate MAS and CPS as follows.
\begin{itemize}
	\item We consider a System as a hierarchy of agents. So, we map agents
		to \emph{systems, sub-systems, or devices}, depending on the
		granularity of the design. For example, a modeler can model a
		specific device as a system not decomposed into sub-systems or devices (and
		the device is considered as an agent).
	\item Agents reason over beliefs (i.e. transform beliefs into other
		beliefs) and each \emph{component of a CPS} (system, sub-system,
		or device) is composed by a \emph{functional architecture} that
		transforms input-beliefs into output-beliefs.\fixnote{Oliviero}{We often times used functional architectures before. Mabybe we should make this connection precise before, then? Or maybe avoid using this CPS-oriented terminology before, and keep it more abstract?}
	\item Components of a CPS have \emph{ports} to exchange information
		with the outer environment (which may be a sub-system),
		similarly to agents. The transfer of information in a CPS is
		defined by \emph{channels}.
	\item The concept of facts is related to the \emph{requirements}
		that describe how the CPS shall behave and
		its \emph{physical architecture}.

@@

\begin{proof}
An input port is, in the $\abftheory$-framework, defined secure as long as the relation
	between the two regions of input assertions $\assertionRegion$ and
	output beliefs $\beliefRegion$ are equal, i.e.
	$\eq{\assertionRegion}{\beliefRegion}$. Therefore, any other relation
	should result in a weakness (related to an insecurity flaw) of that
	input port.  Using RCC5, there exist exactly other $4$ different types
	of relations, one of which is the discrete-from (DR) relation, i.e.
	$\dr{\assertionRegion}{\beliefRegion}$. When two regions are related by
	the DR relations, they have no subregion in common. Let us
	define a function weight $|X|$ such that, for any region $X$, it
	represents the smallest possible cardinality of a (mereo)topological base for
	$X$; where a base is a collection of regions in a (mereo)topology such that
	%every open region\fixnote{Oliviero}{First time we refer to open regions; perhaps briefly introduce them?} can be written as union of elements of that base.
	every region can be written as union of elements of that base. 
	We distinguish between regions that are related to information and regions that
	don't by writing the latter as $\emptyset$ (e.g. a region corresponding to
	an assertion is written $A\neq\emptyset$, while a region corresponding to an empty assertion
	as $A=\emptyset$).
	\begin{enumerate}
		\item If $\eq{\assertionRegion}{\beliefRegion}$ then either
			$\assertionRegion=\beliefRegion=\emptyset$ (no
			communication) or
			$\assertionRegion=\beliefRegion\neq\emptyset$ (forward
			communication).
		\item If $\dr{\assertionRegion}{\beliefRegion}$ then
			$\assertionRegion=\emptyset\oplus\beliefRegion=\emptyset$
			(we call \emph{insert} the former and \emph{full drop}
			the latter case), or
			$\assertionRegion\neq\emptyset\wedge\beliefRegion\neq\emptyset\wedge\assertionRegion\neq\beliefRegion$
			which we call \emph{replace} (i.e. drop and insert).
		\item If $\pp{\assertionRegion}{\beliefRegion}$ then
			$\beliefRegion$ contains and extend $\assertionRegion$
			which we call \emph{insertion}.
		\item If $\ppi{\assertionRegion}{\beliefRegion}$ then
			$\assertionRegion$ contains and extend $\beliefRegion$
			which we call \emph{drop} (or selective drop to stress the difference with the full drop).
		\item If $\po{\assertionRegion}{\beliefRegion}$ then a part of
			the $\assertionRegion$ is contained in the
			$\beliefRegion$ which is a combination of
			\emph{selective drop and generation} which we call \emph{injection}.
	\end{enumerate}
\end{proof}

@@

\begin{definition}{\bf Functional Block and Architecture --}\label{def:funblock}
	A functional block of an agent takes beliefs as  inputs (input-beliefs) and
	returns output-beliefs.  A functional architecture is an
	interconnected system of functional blocks.
\end{definition}\fixnote{Oliviero}{We referred to this notion many times in the esposition, and define it precisely only here. I'd perhaps speak more abstractly until this is introduced.}
The quality of a functional block cannot be determined
by the difference between its inputs and outputs (as we did for
ports and channels), because the behavior of a functional block
cannot be determined in general; since any functional block will have 
its own purpose based on functional requirements. 
Therefore, while the semantics of a port is determined by the relation 
between assertions and beliefs, the semantics of a functional block 
is determined by the relation between facts/requirements and I/O beliefs.
In other words, a functional block is a generic agent with no pre-defined general
behavior (while ports and channels have a pre-defined behavior).
In the following, for the sake of simplicity, 
we use the generic region $\beliefRegion$ to refer to the behavior (i.e.
the beliefs generated by the behavior).

@@

\subsection{Security and Insecurity of a System}\label{sec:formula}
\begin{hypothesis}{\bf System Security Design --}\label{hyp:security}
	A system security design is given by a precise system 
	specification over the physical and functional architectures, that
	uniquely, i.e.  tested against the range of designs possible in the
	$\abftheory$-framework, defines the design built on top of those
	requirements.
\end{hypothesis}

\begin{hypothesis}{\bf System Insecurity Design --}\label{hyp:insecurity}
	If, given a system specification as a collection of
	requirements on the system, there exist
	a non-unique design with respect to those requirements, the number of
	equivalent designs that fulfills the requirements quantitatively
	defines the magnitude of insecurity of a system design.
\end{hypothesis}

\fixnote{Oliviero}{Ok, maybe here my inexperience with systems design is showing, but these hypothesis are not super clear to me. I understand the next part quite well, but the actual hypthesis, not so much. What are the designs possible with respect to the framework? Do you mean all the possible configurations of <rcc(A,B),rcc(A,F),rcc(B,F)> for each agent? What does "testing against" this range of possible designs?}

>>>>

@@

		concept of belief (as epistemological

@@

\emph{beliefs} (internal information considered true by the agent), and
%and the information or 
\emph{assertions} (an exchange of information between agents). % it provides.

@@

%As in \autocite{Steup2020epistemology}, 
%the only objective of information is to
%exchange beliefs between agents and, 
As defined in \autocite{Hintikka1993Information},

@@

beliefs to define the actual behavior of an agent or a system; meaning
that an agent has its own beliefs on top of which it bases its behavior,
the consequences of which, in turn, will update its initial beliefs.
%; i.e. we consider
%behaviors as belief revision systems.

@@

@@

	\item We consider a set of axiomatic \emph{facts} determining how the
		system shall be. Similarly, requirements determine which facts
		must hold when the system is implemented\footnote{We consider facts as
		definitory rules\autocite{Hintikka1993Information} since they
		don't define how a real system is finally implemented but how
		it shall be, trough a series of requirements.  Those
		requirements may not hold, through insecurity, in the
		implemented system.}. Specifically, facts/requirements describe:
		the functional architecture of each agent, and the
		physical/structural (HW/SW) architecture of each subsystem of
		agents and agent within a subsystem.

@@

@@

@@

Similarly to \autocite{Santaca2016abf}, we define agents as a meronomy (an
hierarchy of Part-Whole relations) over the constituents that we previously defined
(assertions, beliefs, and facts), based on a standard definition of mereology,
i.e. based on the definition of parthood relation between \emph{parts}.  Due to
the necessity of considering different relations between parts (as we'll show afterwards)
we extend the mereology to a
mereo-topology \autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
considering the relations in Table~\ref{tab:rcc358}.  For the sake of
readability, we use the term \emph{region} both to refer to a mereological part
and to a topological region.  Our aim is to create a meronomy (hierarchy of
part-whole relations) instead of the taxonomies (categorization based on
discrete sets) such as the one provided in \autocite{NIST2020NVD,MITRE2020CVE}
so that we don't need to rely on a scoring system (such as the CVSS) to assign
a quantitative evaluation of the cybersecurity of each entry. Instead, we want
a precise calculation of the number of insecure configurations of a system to
emerge from the mathematical formulation of our cybersecurity hypothesis.

A mereotopology, as defined e.g. in \autocite{Rachavelpula2017mereotopology},
is a mathematical structure where the basic relation between regions
is the reflexive and symmetric relation \emph{Connects With},
that we use to order a universe of agents
$\agentuniverse$ (see in Table~\ref{tab:rcc358}).  We use the Region Connection
Calculus (RCC), as defined in \autocite{bennettLogics,improvingRCC}, to provide
an axiomatization of the mereo-topological concepts. In its broader definition,
the RCC theory is composed by eight axioms, and is known as RCC8. In the text,
for brevity, we will often focus only on RCC5 without loss of generality. Using
RCC5 instead of RCC8 prevents us from considering tangential connections
between spatial regions. However, tangential connections in RCC8 can be
considered as special cases of the more general spatial relations considered in
RCC5.

@@

\paragraph{Number of different configurations of a system}
The general formula to calculate the number of different types of agents is
$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
between $n$ different regions. Therefore, $r^{\binom{n}{k}}$
expresses the number of permutation of $r$
relations over ${\binom{n}{k}}$ elements with repetitions, 
with ${\binom{n}{k}}$ being the number of
$k$-ary combinations of $n$ regions.
In our case, $\binom{n}{k}=3$ since we consider $3$ regions 
($\assertionRegion,\beliefRegion,\factRegion$), and all the relations
considered in the RCC are binary.  Hence, using RCC5 (with five different
spatial relations) over three regions, we can theoretically define up to 125
different type of agents. However, only 54 of the 125 (as showed in
\autocite{improvingRCC}) combinations are topologically correct with respect to
the definition of the relations of RCC5. Generalizing to all the RCCs:

@@

\paragraph{\Rcc{$\assertionRegion$}{$\beliefRegion$} -- Collaboration} In order
to reason on the relation between assertions and beliefs we first need to
consider that, by definition, assertions are defined as transfer of information
between two agents, e.g., between two agents.  Therefore, as depicted in
Figure~\ref{fig:system-agent}, an agent has two main categories of assertions,
input and output assertions.  Given an agent $a$ and a collection of asserted
predicates $\Phi$,
the input assertions are those received by $a$ from an agent
$s$ acting as a sender, $\rassert{s}{a}{\Phi}$; similarly, output assertions
are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. With a slight abuse
of notation, in the text we simply write $\assertionRegion_{s\rightarrow a}$ when we abstract
away from the content of the assertion (i.e. $\Phi$).
We shall consider
two pairs of regions:%\footnote{``I am not asserting, as Lotze did, that a
%relation between $X$ and $Y$ consists of a quality in $X$ and a quality in $Y$
%-- a view which I regard as quite indefensible. -- I assert that a relation $Z$
%between $X$ and $Y$ involves the existence in $X$ of the quality ``having the
%relation $Z$ to $Y$'' so that a difference of relations always involves a
%difference in quality, and a change of relations always involves a change of
%quality.'' --  Ellis J. McTaggart, The Unreality of
%Time \autocite{Mctaggart1908unreality}}
\begin{itemize}
	\item \Rcc{$\assertionRegion_{s\rightarrow a}$}{$\beliefRegion$},
		where the relation between input-assertions and beliefs
		describes the soundness of the execution of the functional
		architecture w.r.t. input elicitation. With more details, the
		ideal specification of the functional architecture, along with
		the expected inputs, defines the functional behavior of an
		ideal system.  If all the inputs (assertions) are correctly
		handled in the functional specification (beliefs) the
		specification is complete.

@@

We are now in the position to define what a secure system is (with respect to
the $\abftheory$-framework) and, based on that definition, what the cybersecurity risk is and
how to quantify it in a risk matrix. The following definition holds for 
abstract systems defined in the $\abftheory$-framework but will be refined
for CPS afterwards in the paper.
\begin{definition}{\bf Cybersecurity of a System or an Agent --}\label{def:security}
	A secure system is a system where CSE-$1$, CSE-$2$, and CSE-$3$ holds for
	each agent composing the system.
\end{definition}
The ISO 31000 consider risk as the ``effect of uncertainty on objectives'' and
refers both to positive and negative consequences of uncertainty.
Accordingly, we consider risk as follows.
\begin{definition}{\bf Risk --}
The risk is the uncertainty related to the whole space of potential designs
	resulting from a specification in the $\abftheory$-framework.
\end{definition}
The definition of Risk leads to the risk matrix in
Figure~\ref{fig:quantitative}, defined as follows.
\begin{definition}{\bf Risk Matrix --}
	The risk matrix, as summarized in Table~\ref{tab:5com}, is a function of the three relations
	$s=\langle\rcc(\factRegion,\beliefRegion),\rcc(\factRegion,\assertionRegion),\rcc(\beliefRegion,\assertionRegion)\rangle$,
	where the maximum risk is defined by the $DR$ relation between the
	three groups of regions, and the minimum risk by the $EQ$ relation over
	the same regions. In between the two extremes, the granularity of
	possible intermediate configuration is defined by the calculus used
	(RCC5 in our case).
\end{definition}
While a risk matrix is often defined as a function of the likelihood and impact
of attacks (based on quantitative ad-hoc estimation of how likely it is that an
attacker will exploit one or more vulnerabilities and what is the magnitude of
the incidents produced by this exploitation), we suggest that cybersecurity
weaknesses are all equally likely to be exploited if there's a connection
between the weakness and the asset that an attacker wants to impact. When the
whole system is considered an asset, all weaknesses are equally likely to be
exploited.  Therefore, a risk matrix should capture the number of insecure
configurations of a system, rather than predicating over likelihoods of
weaknesses exploitation.

@@

So far, we have reasoned on systems in the abstract, considering
generic Multi-Agent System (MAS); mapping Epistemological concepts
to MAS. We now map MAS to CPS and show how this mapping allows us
to predict cybersecurity weaknesses in a system.

@@

As depicted in Figure~\ref{fig:mashyp}, we relate MAS and CPS as follows.
\begin{itemize}
	\item We consider a System as a hierarchy of agents. So, we map agents
		to \emph{systems, sub-systems, or devices}, depending on the
		granularity of the design. For example, a modeler can model a
		specific device as a system not decomposed into sub-systems or devices (and
		the device is considered as an agent).
	\item Agents reason over beliefs (i.e. transform beliefs into other
		beliefs) and each \emph{component of a CPS} (system, sub-system,
		or device) is composed by a \emph{functional architecture} that
		transforms input-beliefs into output-beliefs.
	\item Components of a CPS have \emph{ports} to exchange information
		with the outer environment (which may be a sub-system),
		similarly to agents. The transfer of information in a CPS is
		defined by \emph{channels}.
	\item The concept of facts is related to the \emph{requirements}
		that describe how the CPS shall behave and
		its \emph{physical architecture}.

@@

\begin{proof}
An input port is, in the $\abftheory$-framework, defined secure as long as the relation
	between the two regions of input assertions $\assertionRegion$ and
	output beliefs $\beliefRegion$ are equal, i.e.
	$\eq{\assertionRegion}{\beliefRegion}$. Therefore, any other relation
	should result in a weakness (related to an insecurity flaw) of that
	input port.  Using RCC5, there exist exactly other $4$ different types
	of relations, one of which is the discrete-from (DR) relation, i.e.
	$\dr{\assertionRegion}{\beliefRegion}$. When two regions are related by
	the DR relations, they have no subregion in common. Let us
	define a function weight $|X|$ such that, for any region $X$, it
	represents the smallest possible cardinality of a (mereo)topological base for
	$X$; where a base is a collection of regions in a (mereo)topology such that
	every region can be written as union of elements of that base. 
	We distinguish between regions that are related to information and regions that
	don't by writing the latter as $\emptyset$ (e.g. a region corresponding to
	an assertion is written $A\neq\emptyset$, while a region corresponding to an empty assertion
	as $A=\emptyset$).
	\begin{enumerate}
		\item If $\eq{\assertionRegion}{\beliefRegion}$ then either
			$\assertionRegion=\beliefRegion=\emptyset$ (no
			communication) or
			$\assertionRegion=\beliefRegion\neq\emptyset$ (forward
			communication).
		\item If $\dr{\assertionRegion}{\beliefRegion}$ then
			$\assertionRegion=\emptyset\oplus\beliefRegion=\emptyset$
			(we call \emph{insert} the former and \emph{full drop}
			the latter case), or
			$\assertionRegion\neq\emptyset\wedge\beliefRegion\neq\emptyset\wedge\assertionRegion\neq\beliefRegion$
			which we call \emph{replace} (i.e. drop and insert).
		\item If $\pp{\assertionRegion}{\beliefRegion}$ then
			$\beliefRegion$ contains and extend $\assertionRegion$
			which we call \emph{insertion}.
		\item If $\ppi{\assertionRegion}{\beliefRegion}$ then
			$\assertionRegion$ contains and extend $\beliefRegion$
			which we call \emph{drop} (or selective drop to stress the difference with the full drop).
		\item If $\po{\assertionRegion}{\beliefRegion}$ then a part of
			the $\assertionRegion$ is contained in the
			$\beliefRegion$ which is a combination of
			\emph{selective drop and generation} which we call \emph{injection}.
	\end{enumerate}
\end{proof}

@@

\begin{definition}{\bf Functional Block and Architecture --}\label{def:funblock}
	A functional block of an agent takes beliefs as  inputs (input-beliefs) and
	returns output-beliefs.  A functional architecture is an
	interconnected system of functional blocks.
\end{definition}
The quality of a functional block cannot be determined
by the difference between its inputs and outputs (as we did for
ports and channels), because the behavior of a functional block
cannot be determined in general; since any functional block will have 
its own purpose based on functional requirements. 
Therefore, while the semantics of a port is determined by the relation 
between assertions and beliefs, the semantics of a functional block 
is determined by the relation between facts/requirements and I/O beliefs.
In other words, a functional block is a generic agent with no pre-defined general
behavior (while ports and channels have a pre-defined behavior).
In the following, for the sake of simplicity, 
we use the generic region $\beliefRegion$ to refer to the behavior (i.e.
the beliefs generated by the behavior).

@@

\subsection{Security and Insecurity of a System}\label{sec:formula}
\begin{hypothesis}{\bf System Security Design --}\label{hyp:security}
	A system security design (in the $\abftheory$-framework) is given by a
	precise system specification over the physical and functional
	architectures that uniquely defines the design built on top of those
	requirements.
\end{hypothesis}

\begin{hypothesis}{\bf System Insecurity Design --}\label{hyp:insecurity}
	If, given a system specification as a collection of requirements, there
	exist a non-unique design with respect to those requirements, the
	number of equivalent designs (that fulfills the requirements)
	quantitatively defines the magnitude of insecurity of a system design.
\end{hypothesis}

Based on these hypothesis, we can formulate the concepts of security and
insecurity (in the $\abftheory$-framework) as mathematical equations.  Let us
consider a CPS $S$, represented as a graph $G=\langle V,E\rangle$ where $V$
represents the set of functional blocks and ports of $S$, and $E\subseteq
V\times V$ is the set of pairs representing the channels and connections (data
flows) between functional blocks. We define $R\subseteq V\times F$, where $F$
is the set of all the requirements of $S$,
and extend $G$ as $G'=\langle V',E'\rangle$ with $E'= E\cup R$ and $V'=V\cup F$. Let $\pi:
E'\rightarrow RCC$ (where $RCC$ is the set of relations in the RCC) be the
total function associating an RCC relation to each edge in $G'$, and $\Pi$ be
the set of all different permutations of RCC relations over $E'$ (i.e.
$\Pi=\{\langle\pi(e_0)=EQ,\ldots,\pi(e_n)=EQ\rangle,\ldots,\langle
\pi(e_0)=DR,\ldots,\pi(e_n)=DR\rangle\}$ where $e_i\in E'$ for $0\leq i \leq n$ and $|E'|=n$). If $\sigma:\Pi\rightarrow\{0,1\}$ is
an evaluation function such that $\sigma(p)=1$ (where $p\in\Pi$) iff the input
configuration is satisfiable with respect to the logical theory defining the
algebraic structure (mereotopology) and constraints of the calculus RCC
(otherwise $\sigma$ returns $0$), we can define
\begin{displaymath}
	I=\sum_{p\in(\Pi\setminus\pi_{eq})} \sigma(p)
\end{displaymath}
where $\pi_{eq}\in\Pi$ is the output of the function $\pi$ that associates only EQ relations
to any $e\in E'$, and $I\in\mathbb{N}$ represents all the insecurity configurations
of the CPS $S$ where, at least, one of the RCC relations isn't EQ.
In other words, we consider $\pi_{eq}$ as the only secure configuration.
